---
title: "Horizon thickness"
author: "Amanda Pennino" modded by Jenny Bower
date: "7/8/2021"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(sf)
library(sp)
library(stringr)
```

Notes: All working, but Latwag and Legacy need addition of calculation of pedon thickness and pedon bottom interface

Ideally you should classify the BC and C/D interfaces and take out all of the extra columns used to calculate things

*Note: currently calculation of Bh horizon thickness is messed up (uses Bhs in addition to Bh)

Add new horizon (horizon_calc) column that is based on color and location

Eventually incorporate spatial things into key
```{r}
AUGERS <- read_csv("data/AugerDescriptions.csv")

AUGERS <- AUGERS %>% 
  rename(
    Site2 = `Site Name`,
    Upper_cm = `Top (cm)`,
    Lower_cm = `Bottom (cm)`,
    Horizon_thickness = `Horizon Depth (cm)`,
    Total_Depth = `Total Depth`,
    Bottom_Interface = `Bottom Interface`
    )

#Remove anything that doesn't have horizon level data! OR! Keep if you're just interested in mapping the HPU
unique(AUGERS$Horizon)
AUGERS <- AUGERS %>% 
  filter(!(Horizon %in% c("E Podzol", "(Bark?) Histosol", "unknown", "Bhs podzol", "Bedrock", "Histosol", "Unknown podzol", "Bedrock", "Dark E Podzol", "Bhs Podzol", "E Podzol\n", "Bedrock Histosol", "Moss", "", "????")))

#Combine major horizon and subdesignations into one column
AUGERS$Horizon2 <- paste(str_trim(AUGERS$Horizon), AUGERS$Suffix, sep = "")
AUGERS$Horizon2 <- gsub('NA', '', AUGERS$Horizon2)

#Combine subwatershed and site into one column
AUGERS$Site <- paste(AUGERS$Subwatershed, AUGERS$Site2, sep = "")

#Remove any blank horizons
AUGERS <- AUGERS %>% 
  filter(!(Horizon == ""))

#Take out subwatershed, HPU, and bottom interface data to add back later after pivots
EXTRA <- AUGERS %>% 
  dplyr::select(Site, HPU, Bottom_Interface, Total_Depth) %>%
  group_by(Site) %>%
  summarise_all(first)

#Select columns to pivot, Take out duplicated rows prior to pivot.
#Need subwatershed otherwise coincident duplicate rows will be removed.
AUGERS <- AUGERS %>% 
  dplyr::select(Site, Horizon2, Horizon_thickness)

AUGERS <- AUGERS %>% mutate(row = row_number()) %>%pivot_wider(names_from = Horizon2, values_from = c(Horizon_thickness)) %>% dplyr::select(-row) %>% group_by(Site) %>% summarise_all(tibble::lst(sum), na.rm = TRUE)

#Add back extra data
AUGERS <- left_join(AUGERS, EXTRA, by = "Site")

#Aggregating horizons for summary stats. e.g., total thickness of all O-horizons
#TAKE NOTE HERE TO WHAT HORIZONS I AM GROUPING INTO WHAT. THIS CAN BE CHANGED.
#dimnames(AUGERS)
Ofun <- function(x){
  x <- cbind(x, allO_thickness = rowSums(x[,grep("O",colnames(x))], na.rm = TRUE))
  return(x$allO_thickness)}
Efun <- function(x){
  x <- cbind(x, allE_thickness = rowSums(x[,grep("E",colnames(x))], na.rm = TRUE))
  return(x$allE_thickness)}
Bhsfun <- function(x){
  x <- cbind(x, allBhs_thickness = rowSums(x[,grep("Bhs",colnames(x))], na.rm = TRUE))
  return(x$allBhs_thickness)}
Bsfun <- function(x){
  x <- cbind(x, allBs_thickness = rowSums(x[,grep("Bs",colnames(x))], na.rm = TRUE))
  return(x$allBs_thickness)}
Bhfun <- function(x){
  x <- cbind(x, allBh_thickness = rowSums(x[,grep("Bh[^s]",colnames(x))], na.rm = TRUE))
  return(x$allBh_thickness)}
Bgfun <- function(x){
  x <- cbind(x, allBg_thickness = rowSums(x[,grep("Bg",colnames(x))], na.rm = TRUE))
  return(x$allBg_thickness)}
Bwfun <- function(x){
  x <- cbind(x, allBw_thickness = rowSums(x[,grep("Bw",colnames(x))], na.rm = TRUE))
  return(x$allBw_thickness)}
BCfun <- function(x){
  x <- cbind(x, allBC_thickness = rowSums(x[,grep("BC",colnames(x))], na.rm = TRUE))
  return(x$allBC_thickness)}
Cfun <- function(x){
  x <- cbind(x, allC_thickness = rowSums(x[,grep("[^B]C",colnames(x))], na.rm = TRUE))
  return(x$allC_thickness)}

#Apply functions
allO <- Ofun(AUGERS)
allE <- Efun(AUGERS)
allBhs <- Bhsfun(AUGERS)
allBs <- Bsfun(AUGERS)
allBh <- Bhfun(AUGERS)

#Add back to data frame
AUGERS$allO_thickness <- allO
AUGERS$allE_thickness <- allE
AUGERS$allBhs_thickness <- allBhs
AUGERS$allBs_thickness <- allBs
AUGERS$allBh_thickness <- allBh

#Add locations
#This makes sure to join based on subwatershed PLUS site
LOCATIONS_AUGERS <- read_sf(dsn = "data/Shapefiles", layer = "auger_locations_allsub")
LOCATIONS_AUGERS <- LOCATIONS_AUGERS %>% rename(Site2 = Site)
LOCATIONS_AUGERS$Site <- paste0(LOCATIONS_AUGERS$Sub, LOCATIONS_AUGERS$Site2)
AUGERS <- left_join(AUGERS, LOCATIONS_AUGERS, by = "Site")
AUGERS_shp <- AUGERS[,c("Site", "Total_Depth", "allO_thickness", "allE_thickness", "allBhs_thickness", "allBs_thickness", "allBh_thickness", "x", "y", "geometry")]
write.csv(AUGERS, "augers.csv")

#Exports auger file
sf::st_write(AUGERS_shp, "data/Shapefiles/HorizonThickness_Augeronly.shp", delete_layer=TRUE)

```

LATWAG soil pits (Jenny's, Steph's, PFMs)
```{r}
#update as data gets added to the database
#SET IT TO YOUR OWN DIRECTORY
LATWAG <- read_excel("~/Documents/uvm/phd/research/data/lab/Soil_Analyses/data/raw/hb_pedons_analysis_0426.xlsx", sheet = "horizon data")

#consider changing to pedons_analysis database
#need to calculate lower depth & add in bottom interface thru a summarize function (sum thickness, run a last on horizon)
LATWAG <- LATWAG %>% 
  rename(
    Site = pedon,
    Upper_cm = `top cm`,
    Lower_cm = `base cm`,
    Horizon_thickness = thick
    )

LATWAG$Horizon2 <- LATWAG$horizon

LATWAG <- LATWAG %>%
  dplyr::select(Site, Horizon2, Horizon_thickness)

LATWAG <- LATWAG %>% mutate(row = row_number()) %>%pivot_wider(names_from = Horizon2, values_from = c(Horizon_thickness)) %>% dplyr::select(-row) %>% group_by(Site) %>% summarise_all(tibble::lst(sum), na.rm = TRUE)

#Apply functions
allO <- Ofun(LATWAG)
allE <- Efun(LATWAG)
allBhs <- Bhsfun(LATWAG)
allBs <- Bsfun(LATWAG)
allBh <- Bhfun(LATWAG)

#Add back to data frame
LATWAG$allO_thickness <- allO
LATWAG$allE_thickness <- allE
LATWAG$allBhs_thickness <- allBhs
LATWAG$allBs_thickness <- allBs
LATWAG$allBh_thickness <- allBh

LOCATIONS_LATWAG <- read_excel("~/Documents/uvm/phd/research/data/lab/Soil_Analyses/data/raw/hb_pedons_analysis_0426.xlsx", sheet = "pedon data")

#exclude one pit missing coordinates
LOCATIONS_LATWAG <- LOCATIONS_LATWAG %>% filter(!(`Pedon ID` == "42_4_V4.2"))
LOCATIONS_LATWAG <- st_as_sf(LOCATIONS_LATWAG, coords = c("Easting", "Northing"))
LOCATIONS_LATWAG <- LOCATIONS_LATWAG %>% rename(Site = `Pedon ID`)
LOCATIONS_LATWAG <- LOCATIONS_LATWAG %>% dplyr::select(Site, HPU, geometry)
LATWAG <- left_join(LATWAG, LOCATIONS_LATWAG, by = "Site")

sf::st_write(LATWAG, "data/Shapefiles/HorizonThickness_Latwagonly.shp", delete_layer=TRUE)

```

Legacy data from Scott
```{r}
#Load legacy data
LEGACY_pedons <- read_excel("data/HBEF_pedons_20210603.xlsx")
LEGACY_horizons <- read_excel("data/HBEF_soil_horizons_20210603.xlsx")
LEGACY_horizons <- subset(LEGACY_horizons, select=-c(frags))

#Filter only pedons in WS3
LEGACY_pedons <- st_as_sf(LEGACY_pedons, coords = c("easting", "northing"))
st_crs(LEGACY_pedons) <- 26919
WS3 <- read_sf(dsn = "data/Shapefiles", layer = "ws3")
LEGACY_coords <- LEGACY_pedons[WS3, ]
LOCATIONS_LEGACY <- LEGACY_coords %>% filter(!(project == "GRADJAB")) 
LOCATIONS_LEGACY <- st_set_geometry(LOCATIONS_LEGACY, NULL)

#Many-to-one join
LOCATIONS_LEGACY <- merge(LOCATIONS_LEGACY, LEGACY_horizons, by = c("pedon", "pedon"))

#Rename variables
LOCATIONS_LEGACY <- LOCATIONS_LEGACY %>% 
  rename(
    Site = pedon,
    Upper_cm = `top cm`,
    Lower_cm = `base cm`,
    Horizon2 = horizon
    )

#Calc horizon thickness
LOCATIONS_LEGACY$Horizon_thickness <- LOCATIONS_LEGACY$Lower_cm - LOCATIONS_LEGACY$Upper_cm

#Convert back to data frame and select relevant columns
LEGACY <- as_tibble(LOCATIONS_LEGACY)

LEGACY <- LEGACY %>%
  dplyr::select(Site, Horizon2, Horizon_thickness)

LEGACY <- LEGACY %>% mutate(row = row_number()) %>% pivot_wider(names_from = Horizon2, values_from = c(Horizon_thickness)) %>% dplyr::select(-row) %>% group_by(Site) %>% summarise_all(tibble::lst(sum), na.rm = TRUE)

#Apply functions
allO <- Ofun(LEGACY)
allE <- Efun(LEGACY)
allBhs <- Bhsfun(LEGACY)
allBs <- Bsfun(LEGACY)
allBh <- Bhfun(LEGACY)

#Add back to data frame
LEGACY$allO_thickness <- allO
LEGACY$allE_thickness <- allE
LEGACY$allBhs_thickness <- allBhs
LEGACY$allBs_thickness <- allBs
LEGACY$allBh_thickness <- allBh

#This is where I'm currently stuck
LEGACY_coords <- st_as_sf(LEGACY_coords, coords = c("Easting", "Northing"))
LEGACY_coords <- LEGACY_coords %>% rename(Site = pedon, HPU = hpu)
LEGACY_coords <- LEGACY_coords %>% dplyr::select(Site, HPU, geometry)

LEGACY <- left_join(LEGACY, LEGACY_coords, by = "Site")
sf::st_write(LEGACY, "data/Shapefiles/HorizonThickness_Legacyonly.shp", delete_layer=TRUE)
```

Combine datasets, fix differences in HPU designations, and export the shapefile.
```{r}
#Combine all three datasets
X <- bind_rows(AUGERS, LATWAG, LEGACY)

#Fix differences in HPU designation
X$HPU[X$HPU == "T"] <- "Typical"
X$HPU[X$HPU == "Bi"] <- "Bimodal"

#Select only needed fields
X <- select(X, c('allO_thickness', 'allE_thickness', 'allBhs_thickness', 'allBs_thickness', 'allBh_thickness','HPU','Site','x','y','geometry'))

sf::st_write(X, "data/Shapefiles/HorizonThickness_all.shp", driver = "ESRI Shapefile", delete_layer=TRUE)

```

Determine HPU
```{r}
#Bhs Podzol part of script needs tuning to account for landscape position; Bimodal needs tuning to account for Bh being under Bs or Bh

X <- cbind(X, HPU2 = '')
X_shp <- read_sf(dsn = "data/Shapefiles", layer = "HorizonThickness_all")
BOSS <- read_sf(dsn = "data/Shapefiles", layer = "BOSS_only")
X_in_BOSS <- st_join(X_shp, BOSS, join = st_within)
X_in_BOSS <- subset(X_in_BOSS, !is.na(X_in_BOSS$DN))

for(a in seq(1:nrow(X))){
  if(sum(X$allBhs_thickness[a], X$allBs_thickness[a], X$allBh_thickness[a]) == 0 && (X$allO_thickness[a] >= 15 | X$allE_thickness[a] <= 15)){
    X$HPU2[a] <- 'Bedrock Histosol'
  } else if((X$allBs_thickness[a] > 0 && X$allBh_thickness[a] > 0) | (X$allBh_thickness[a] > 0 && (X$allBh_thickness[a] < X$allBhs_thickness[a]))){
    X$HPU2[a] <- 'Bimodal Podzol'
  } else if(X$allBs_thickness[a] == 0 && X$allBh_thickness[a] > 0 && (X$allBhs_thickness[a] < X$allBh_thickness[a])){
    X$HPU2[a] <- 'Bh Podzol'
  } else if(X$allE_thickness[a] > 15 | (X$allE_thickness[a] > sum(X$allBhs_thickness[a], X$allBs_thickness[a], X$allBh_thickness[a]))){
    X$HPU2[a] <- 'E Podzol'
  } else if(X$Site[a] %in% X_in_BOSS$Site && (X$allE_thickness[a] == 0 | (X$allBhs_thickness[a] > X$allE_thickness[a]))){
    X$HPU2[a] <- 'Bhs Podzol'
  } else if(X$allBhs_thickness[a] > 0 | X$allBs_thickness[a] > 0){X$HPU2[a] <- 'Typical Podzol'
  } else{
    X$HPU2[a] <- 'Unclassified'
  }
}

#Export shapefile
sf::st_write(X, "data/Shapefiles/HorizonThickness_all.shp", driver = "ESRI Shapefile", delete_layer=TRUE)

```